Global code motion of congruent instructions

### 2017-11-11 cgo/paper.tex
formatting

We have implemented the pass in \LLVM{}. It is based on Global Value Numbering
infrastructure available in \LLVM{}. The experimental results show an average
saving of 1\% on the total compilation time on \SPEC{}. \GCM{} enables more
inlining and exposes more loop invariant code motion opportunities in the
majority of benchmarks. We have also seen execution time improvements in a few
of \SPEC{} benchmarks viz. mcf and sjeng, moreover, register spills reduced by
2\% on the \SPEC{} benchmarks suite when compiled for \xlinux{}. \GCM{} is an
optimistic algorithm in the sense that it considers all congruent instructions
in a function as potential candidates. We also formalize why register pressure
reduces, and how \GCM{} increases instruction level parallelism thereby enabling
more \ooo{} execution on modern architectures.
\end{abstract}

@@

So we introduce a data structure ($\chi$ nodes) to keep track of values flowing out
of a basic block with more than once successors. We only do this for values with
multiple occurrences in the function as they are the potentially hoistable
candidates. This approach allows us to hoist instructions to a basic block with
more than two successors. The $\chi$ node in a basic block $B$ has CHIArgs with
three entries: a) the instruction (I) CHIArg refers to, b) the basic block $Dest$
which is post-dominated by the basic block containing I, and c) the GVN of I.

@@

When SSAPRE inserts $\Phi$ (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a $\Phi$ has a missing entry,
that means the expression is partially available.  We apply a similar concept
for hoisting expression by working on inverted graph. We insert a $\chi$, to a
basic block with multiple successors. Then for each instruction, we register its
flow of value to its corresponding $\chi$ which can be found in its PDF. For
conciseness we only register values with multiple occurrences. Finally, we start
walking the CFG and remove the $\chi$ if it has a missing entry (see
Figure~\cite{fig:chi-intro2}), that means the expression is not anticipable in
the basic block having $\chi$. For the remaining $\chi$s anticipabilty of `values'
is ensured by construction.

>>>>

We have implemented the pass in \LLVM{}. It is based on Global Value Numbering
infrastructure available in \LLVM{}. The experimental results show an average
saving of 1\% on the total compilation time on \SPEC{}. \GCM{} enables more
inlining and exposes more loop invariant code motion opportunities in the
majority of benchmarks. We have also seen execution time improvements in a few
of \SPEC{} benchmarks viz. mcf and sjeng, moreover, register spills reduced by
2\% on the \SPEC{} benchmarks suite when compiled for \xlinux{}. \GCM{} is an
optimistic algorithm in the sense that it considers all congruent instructions
in a function as potential candidates. We also formalize why register pressure
reduces as a result of \gcm{}, and how \gcm{} increases instruction level
parallelism thereby enabling more \ooo{} execution on modern architectures.
\end{abstract}

@@

So we introduce a data structure ($\chi$ nodes) to keep track of values flowing out
of a basic block with more than one successor. We only do this for values with
multiple occurrences in the function as they are the potentially hoistable
candidates. This approach allows us to hoist instructions to a basic block with
more than two successors. The $\chi$ node in a basic block $B$ has CHIArgs with
three entries: a) the instruction (I) CHIArg refers to, b) the basic block $Dest$
which is post-dominated by the basic block containing I, and c) the GVN of I.

@@

When SSAPRE inserts $\Phi$ (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a $\Phi$ has a missing entry,
that means the expression is partially available.  We apply a similar concept
for hoisting expression by working on inverted graph. We insert a $\chi$, to a
basic block with multiple successors. Then for each instruction, we register its
flow of value to its corresponding $\chi$ which can be found in its PDF. For
conciseness we only register values with multiple occurrences. Finally, we start
walking the CFG and remove the $\chi$ if it has a missing entry (see
Figure~\ref{fig:chi-intro2}), that means the expression is not anticipable in
the basic block having $\chi$. For the remaining $\chi$s anticipabilty of `values'
is ensured by construction.
